---
title: "ACMA320Assignment1"
author: "Timothy Zintel"
date: "25/01/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T )
```

**Assignment Parameters**
The full assignment can be viewed in the attached file "ACMA320_asmt1_2020.pdf"

**Assumptions** 
 Mortality follows Makeham's Law with parameters A, B, and c.
 
 That is, the force of mortality at x = A+Bc^x
 
##Initialization and libraries

```{r}
#install.packages("tidyverse")
library("tidyverse")
set.seed(340)
eps=0.0001
```



##Creating the MakehamSurv function
```{r}
MakehamSurv=function(A,B,c,x,t){
  if(c==1){
    return("Ln(C)=0")
  }
  accid=exp((-1)*A*t)
  ans=exp((-1)*A*t)*(exp(((-1)*(B)/(log(c)))*(c^x)*((c^t)-1)))
  return(ans)
}
```

Here is a test for the function
```{r}
MakehamSurv(0,0.0005,1.05,0,25)
```






##Creating the tpxFiller function
```{r}
tpxFiller=function(A,B,c,x,w,inc){
  result <- matrix(data = NA, nrow= ceiling((w-x)/inc), ncol=2)
  i=1 ##index for loop
  while(i<=nrow(result)){
    result[i]<- inc*i
    if(i*inc>w){
      result[i]<-w
    }
    result[i,2]<- MakehamSurv(A,B,c,x,x+result[i])
    i=i+1
  }
  return(result)
}
```

Testing tpxFiller
```{r}
tpxFiller(0,0.0005,1.05,0,99,7.5)
```



##Creating the pmf function
The u|tqx is tpx-(t+uPx)
```{r}
pmf=function(A,B,c,x,w){
  i=0
  final <- vector(mode="double",length = 0)
  while(i<=(w-x-1)){
    tempvec <- (MakehamSurv(A,B,c,x,i)-MakehamSurv(A,B,c,x,i+1))
    final <- cbind(final,tempvec)
    i=i+1
  }
  return(final)
  
}
```
Testing the pmf function
```{r}
pmf(0,0.0005,1.05,0,100)
```



## Answering the first question
A=0.0005 B=0.0006 c=1.055 w=120

part(a)
```{r}
A= 0.0005
B=0.0006
c=1.055
w=120
tp35g = tpxFiller(A,B,c,35,w,1)
plot(tp35g,type = 'l')
```

part(b)
```{r}
pmf35= pmf(A,B,c,35,w)
plot(seq(from = 0,to = w-35-1),pmf35,type = "h")
```


## Answering the second question
Part (a) summing the Makeham Surv Function
```{r}
curtexp=0
i=1
x=35
n=10
while(i<=n){
  curtexp= curtexp+ MakehamSurv(A,B,c,35,i)
  i=i+1
}
print(curtexp)
```
Part (b) Finding the standard deviation
```{r}
i=1
sum=0
while(i<=n){
    sum=sum+(i*MakehamSurv(A,B,c,35,i))
    i=1+i
  }
ans=(((2*(sum))-curtexp)-(curtexp*curtexp))
print(sqrt(ans))
```
## Answering the third question
A) N year temporary complete expectation of life
```{r}
stop=0
 j=1
 lastesti=1+eps
 estimate=0
   inc=0.75
   while(abs(estimate-lastesti)>eps){
   lastesti = estimate
   i=1
   integral=0
   while((i*inc)<=n){
     integral =integral+(inc*(((MakehamSurv(A,B,c,x,inc*i))+(MakehamSurv(A,B,c,x,inc*(i+1))))/2))
     i=i+1
   }
     estimate= integral
     j=j+1
     inc=inc/2 ##To increase accuracy
     stop=stop+1
     if(stop>11){
       break ##In case of divergence
     }
  }
  compexp=estimate
print(compexp)
  
```

B) Using the second moment to find the variance
```{r}
stop=0
 j=1
 lastesti=1+eps
 estimate=0
  inc=0.75
   while(abs(estimate-lastesti)>eps){ ##Checking to see within 3 decimal places
   lastesti = estimate
   i=1
   integral=0
   while((i*inc)<=n){
     integral =integral+inc*(inc*(((MakehamSurv(A,B,c,x,inc*i))+(MakehamSurv(A,B,c,x,inc*(i+1))))/2))
     i=i+1 ##Added in the t multiplier to the tpx
   }
     estimate= integral
     j=j+1
     inc=inc/2 ##To increase accuracy
     stop=stop+1
     if(stop>11){
       break ##In case of divergence
     }
  }
  moment2=2*estimate
print(moment2)
txvar= moment2-(compexp*compexp)
print("The standard deviation is")
print(sqrt(abs(txvar)))
```

